---
title: "HW 2"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.align = "center", warning = FALSE, message = FALSE, R.options=list(max.print=100))

# Edit the file starting below
```

## Enter your name and EID here

**This homework is due on February 3, 2023 by 11:59pm. You will submit a pdf file on Gradescope.**

*For all questions, include the R code that you used to find your answer (show R chunk and outputs). Answers without supporting code will not receive credit (unless no code was required). Outputs without comments will not receive credit either: Write full sentences to describe your findings.*

------------------------------------------------------------------------

### Question 1: (2 pts)

##### The dataset `ChickWeight` contains information about the weights (in grams) of chicks on four different diets over time (measured at 2-day intervals) as the result of an experiment. The first few observations are listed below.

```{r}
# Load the tidyverse (including ggplot2)
library(tidyverse)

# Save dataset as a tidyverse dataframe (tibble)
ChickWeight <- as_tibble(ChickWeight)

# Visuaize the first ten rows of the dataset
head(ChickWeight, 8)
```

Use some combination of `dplyr` functions to answer the following questions: 

- How many distinct chicks are there?  
- How many distinct time points are there?  
- How many distinct diet conditions are there?  
- How many data points are there per diet condition? 

HINT: The `distinct()` function returns the unique elements of a set of values.


```{r}
## add your code here (and edit this comment to make it descriptive)
```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 2: (2 pts)

##### Using `ggplot2` functions, create a simple scatterplot showing chick `weight` (on the y-axis) as a function of `Time`. Label the axes including the units of the variables and give the plot a title. How does chick `weight` change over `Time`?

```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 3: (2 pts)

Building upon the previous plot, map `Chick` to an aesthetic that assigns a color to each distinct chick’s data points. Add lines that connect each chick’s points together with `geom_line()`. Finally, remove the legend. Do all chicks seem to gain weight in the same manner? Why/Why not?


```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 4: (2 pts)

Continue modifying the same plot by 

* removing the color from the points only
* make all of the points white 
* leave the lines colored by chick
* Put the points *on top of* the lines 

On which day was the last value of the chicks' weight recorded?

```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 5: (2 pts)

Now, facet this plot by diet. Can you tell from this new plot which diet results in greater weight? Explain.

```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 6: (2 pts)

Go back to your plot from question 2 and fit a *linear regression line* (using `lm`) to the chicks in each diet with `geom_smooth()`. There should be 4 separate regression lines, one for each diet, each a separate color. Can you see more clearly which diet results in greater weight? Explain.

```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 7: (2 pts)

The following questions will make use of a dataset on daily temperature and mortality in New York City, New York for the years 1987--2005. The dataset file is included in the homework directory (`data.csv.gz`). 

NOTE: The file is in a compressed format (the `.gz` extension) but it can be read directly into R using `read_csv()`.

The dataset is a data frame with three columns:

* `date`: the date of the observation in YYYY-MM-DD format
* `temp`: the temperature in degrees Fahrenheit
* `death`: the number of deaths from all non-accidental causes on that day

You can read in the dataset using the following code.

```{r}
## Read in temperature and mortality dataset 
dat <- read_csv("data.csv.gz")
```

Use tabulation techniques to summarize the relationship between temperature and mortality. Use `cut_number()` to create a new variable that divides the temperature variable into 5 categories. Then compute the mean of `death` in each temperature category using `group_by()` and `summarize()`. 

Describe what pattern you see with the output. From the output, how would you characterize the relationship between temperature and mortality?

```{r}
## add your code here (and edit this comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 8: (2 pts)

Use `ggplot2` functions to make a scatterplot of temperature on the x-axis and mortality on the y-axis. Add a linear regression line to the plot using `geom_smooth(method = "lm")` and add a flexible smoother to your scatterplot using `geom_smooth()`. Make sure to label your axes informatively and provide a title for the plot. 

Describe the general trends that you see in the scatterplot. From the scatterplot output, how would you characterize the relationship between temperature and mortality? In particular, what is happening with the data in the range of temperature greater than 80 degrees? 

```{r}
## add your code here (and edit this comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 9: (2 pts)

The tabular output using 5 categories of temperature and the scatterplot show different summaries of the data. How could you modify the tabular summary of the data to better match the pattern seen in the scatterplot?

**Your answer goes here. Write sentences in bold.**


From what you have seen, what would you hypothesize is the true relationship between daily temperature and mortality in New York City?

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Formatting: (2 pts)

##### Comment your code, write full sentences, and knit your file!

------------------------------------------------------------------------