---
title: "HW 2"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.align = "center", warning = FALSE, message = FALSE, R.options=list(max.print=100))

# Edit the file starting below
```

## Enter your name and EID here

**This homework is due on February 3, 2023 by 11:59pm. You will submit a pdf file on Gradescope.**

*For all questions, include the R code that you used to find your answer (show R chunk and outputs). Answers without supporting code will not receive credit (unless no code was required). Outputs without comments will not receive credit either: Write full sentences to describe your findings.*

------------------------------------------------------------------------

### Question 1: (2 pts)

##### The dataset `ChickWeight` contains information about the weights (in grams) of chicks on four different diets over time (measured at 2-day intervals) as the result of an experiment. The first few observations are listed below.

```{r}
# Load the tidyverse (including ggplot2)
library(tidyverse)

# Save dataset as a tidyverse dataframe (tibble)
ChickWeight <- as_tibble(ChickWeight)

# Visuaize the first ten rows of the dataset
head(ChickWeight, 8)
```

Use some combination of `dplyr` functions to answer the following questions: 

- How many distinct chicks are there?  
- How many distinct time points are there?  
- How many distinct diet conditions are there?  
- How many data points are there per diet condition? 

HINT: The `distinct()` function returns the unique elements of a set of values.


```{r}
## add your code here (and edit this comment to make it descriptive)
```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 2: (2 pts)

##### Using `ggplot2` functions, create a simple scatterplot showing chick `weight` (on the y-axis) as a function of `Time`. Label the axes including the units of the variables and give the plot a title. How does chick `weight` change over `Time`?

```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 3: (2 pts)

Building upon the previous plot, map `Chick` to an aesthetic that assigns a color to each distinct chick’s data points. Add lines that connect each chick’s points together with `geom_line()`. Finally, remove the legend. Do all chicks seem to gain weight in the same manner? Why/Why not?


```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 4: (2 pts)

Continue modifying the same plot by 

* removing the color from the points only
* make all of the points white 
* leave the lines colored by chick
* Put the points *on top of* the lines 

On which day was the last value of the chicks' weight recorded?

```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 5: (2 pts)

Now, facet this plot by diet. Can you tell from this new plot which diet results in greater weight? Explain.

```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 6: (2 pts)

Go back to your plot from question 2 and fit a *linear regression line* (using `lm`) to the chicks in each diet with `geom_smooth()`. There should be 4 separate regression lines, one for each diet, each a separate color. Can you see more clearly which diet results in greater weight? Explain.

```{r}
## add your code here (and edit this comment to make it descriptive)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 7: (2 pts)

A scatterplot might not be the best way to visualize this data: it calls attention to the relationship between weight and time, but it can be hard to see the differences between diets. A more traditional approach for exploring the effect of diet would be to construct a barplot representing group means with standard error bars showing +/- 1 standard error.

Create a plot using `geom_bar()` where each bar's height corresponds to the average chick weight for each of the four diet conditions. Rename the y-axis to include units (e.g., with scale_y_continuous(`name=`...)) and make the major tick marks go from 0 to 150 by 25 (e.g., with scale_y_continuous(`breaks=`...)). Which diet has the highest mean `weight`? 

```{r}
## add your code here (and edit this comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 8: (2 pts)

Add error bars showing + or - 1 SE using `geom_errorbar(stat = "summary")`. Make the error-bars skinnier by adding a `width =`**0.5** argument. Color the bars (not the error bars, but the barplot bars) by diet and change from the default color scheme using a `scale_fill_` or a `scale_color_`. diet seems to have the most variation in `weight`? The least variation?

```{r}
## add your code here (and edit this comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 9: (2 pts)

Take your code from question 8 and replace `geom_bar()` with `geom_point()`. Remove the `breaks=` argument from `scale_y_continuous`. Make the points larger and color them all red. Put them *on top of* the error bars. Does the mean chick weight seem to differ based on the diet? *I am not asking to conduct hypothesis testing but informally state if they seem to differ and if so, how.*

```{r}
## add your code here (and edit this comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Formatting: (2 pts)

##### Comment your code, write full sentences, and knit your file!

------------------------------------------------------------------------